---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Benefit Only - Induction off
The following includes the R scripts that we used to generate our graphs for the data where the potential benefit to the hosts of lysogenic phage was turned on, but induction was turned off.

## Gather Settings and Treatments

General settings
```{r}
pro_folder <- "Data/BenefitOnly/" #Folder where datafiles are 
pro_treatments <- c("PLR", "PIV") #Names of the treatments being tested - should match what is in filenames
```

Stacked histogram settings
```{r}
lysis_separate_by <- "PIV" #Facet wrap lysis chance stacked histogram by Prophage Loss Rate
#How lysis chance stacked histogram bins should be collapsed and renamed
lysis_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Highly lysogenic)",
                          Hist_0.1 = "0 to 0.2 (Highly lysogenic)",
                          Hist_0.2 = "0.2 to 0.4 (Moderately lysogenic)",
                          Hist_0.3 = "0.2 to 0.4 (Moderately lysogenic)",
                          Hist_0.4 = "0.4 to 0.6 (Temperate)",
                          Hist_0.5 = "0.4 to 0.6 (Temperate)",
                          Hist_0.6 = "0.6 to 0.8 (Moderately lytic)",
                          Hist_0.7 = "0.6 to 0.8 (Moderately lytic)",
                          Hist_0.8 = "0.8 to 1.0 (Highly lytic)",
                          Hist_0.9 = "0.8 to 1.0 (Highly lytic)")

incorp_diff_separate_by <- "PIV"
incorp_diff_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Highly beneficial)",
                          Hist_0.1 = "0 to 0.2 (Highly beneficial)",
                          Hist_0.2 = "0.2 to 0.4 (Moderately beneficial)",
                          Hist_0.3 = "0.2 to 0.4 (Moderately beneficial)",
                          Hist_0.4 = "0.4 to 0.6 (Neutral)",
                          Hist_0.5 = "0.4 to 0.6 (Neutral)",
                          Hist_0.6 = "0.6 to 0.8 (Moderately detrimental)",
                          Hist_0.7 = "0.6 to 0.8 (Moderately detrimental)",
                          Hist_0.8 = "0.8 to 1.0 (Highly detrimental)",
                          Hist_0.9 = "0.8 to 1.0 (Highly detrimental)")
```

## Collect and Munge Data

Gather filenames
```{r}
pro_all_filenames <- list.files(pro_folder)

pro_freeliving_filenames <- str_subset(pro_all_filenames, "FreeLivingSyms")
pro_hostval_filenames <- str_subset(pro_all_filenames, "HostVals")
pro_incvaldiffs_filenames <- str_subset(pro_all_filenames, "IncValDifferences")
pro_lysischance_filenames <- str_subset(pro_all_filenames, "LysisChance")
pro_phagevals_filenames <- str_subset(pro_all_filenames, "SymVals")
```

Combine time series data for all subsets of datafiles
```{r, results='hide', message=FALSE}
pro_freeliving <- combine_time_data(pro_freeliving_filenames, pro_folder, pro_treatments)
pro_hostvals <- combine_time_data(pro_hostval_filenames, pro_folder, pro_treatments)
pro_incvalsdiffs <- combine_time_data(pro_incvaldiffs_filenames, pro_folder, pro_treatments)
pro_lysischances <- combine_time_data(pro_lysischance_filenames, pro_folder, pro_treatments)
pro_phagevals <- combine_time_data(pro_phagevals_filenames, pro_folder, pro_treatments)
```

Rearrange time series data into stacked histogram data
```{r}
pro_lysis_histdata <- combine_histogram_data(pro_lysischances, 
                                         lysis_separate_by, 
                                         lysis_histogram_bins)

pro_incorp_diff_histdata <- combine_histogram_data(pro_incvalsdiffs,
                                               incorp_diff_separate_by,
                                               incorp_diff_histogram_bins)
```

Extract information about average genome values at the end of the runs
```{r}
pro_final_freeliving <- last_update(pro_freeliving)
pro_final_hostvals <- last_update(pro_hostvals)
pro_final_incvalsdiffs <- last_update(pro_incvalsdiffs)
pro_final_lysischances <- last_update(pro_lysischances)
pro_final_phagevals <- last_update(pro_phagevals)
```

## Analyze Data
### Summary Statistics
```{r}
pro_lysis_summary_stats <- pro_final_lysischances %>%
  group_by(PLR, PIV) %>%
  summarise(final_chance_of_lysis = mean(mean_lysischance*100))

pro_lysis_summary_stats
```

```{r}
pro_incvals_summary_stats <- pro_final_incvalsdiffs %>%
  filter(!is.na(mean_incval_difference)) %>%
  group_by(PLR, PIV) %>%
  summarise(final_incval_diffs = mean(mean_incval_difference))

pro_incvals_summary_stats
```

### Wilcox tests
PLR of 5% - compare dormant phage to harmful phage
```{r}
noInd_PLR5 <- subset(plr_final_lysischances, PLR == "0.05")
pro_PLR5_PIV0 <- subset(subset(pro_final_lysischances, PLR == "0.05"), PIV == "0")

wilcox.test(noInd_PLR5$mean_lysischance, pro_PLR5_PIV0$mean_lysischance)$p.value * bonferroni
```

PLR of 2.5% - compare dormant phage to harmful phage
```{r}
noInd_PLR2.5 <- subset(plr_final_lysischances, PLR == "0.025")
pro_PLR2.5_PIV0 <- subset(subset(pro_final_lysischances, PLR == "0.025"), PIV == "0")

wilcox.test(noInd_PLR2.5$mean_lysischance, pro_PLR2.5_PIV0$mean_lysischance)$p.value * bonferroni
```

PLR of 5% - compare dormant phage to beneficial phage
```{r}
pro_PLR5_PIV1 <- subset(subset(pro_final_lysischances, PLR == "0.05"), PIV == "1")

wilcox.test(noInd_PLR5$mean_lysischance, pro_PLR5_PIV1$mean_lysischance)$p.value * bonferroni
```

PLR of 2.5% - compare dormant phage to beneficial phage
```{r}
pro_PLR2.5_PIV1 <- subset(subset(pro_final_lysischances, PLR == "0.025"), PIV == "1")

wilcox.test(noInd_PLR2.5$mean_lysischance, pro_PLR2.5_PIV1$mean_lysischance)$p.value * bonferroni
```

## Generate Graphs
### Host graphs
Host count
```{r}
hostcount_plot <- ggplot(data=pro_hostvals,
                         aes(x=update, y=count, 
                             group=PLR, colour=PLR)) + 
  ylab("Host Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(pro_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_hostvals$PLR)+2))

hostcount_plot + facet_wrap(~PIV)
```

Host cfu count
```{r}
hostcount_cfu_plot <- ggplot(data=pro_hostvals, aes(x=update, y=cfu_count, 
                                                group=PLR, colour=PLR)) + 
  ylab("Host CFU Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(pro_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_hostvals$PLR)+2))

hostcount_cfu_plot + facet_wrap(~PIV)
```

Host uninfected count
```{r}
host_uninfected_plot <- ggplot(data=pro_hostvals, aes(x=update, y=uninfected_host_count, 
                                                  group=PLR, colour=PLR)) + 
  ylab("Uninfected Host Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(pro_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_hostvals$PLR)+2))

host_uninfected_plot + facet_wrap(~PIV)
```

Host int vals
```{r}
hostvals_plot <- ggplot(data=pro_hostvals, aes(x=update, y=mean_intval, 
                                           group=PLR, colour=PLR)) + 
  ylab("Mean Host Interaction Value") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(pro_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_hostvals$PLR)+2))

hostvals_plot + facet_wrap(~PIV)
```

### Phage graphs

Phage count
```{r}
phagecount_plot <- ggplot(data=pro_lysischances,
                          aes(x=update, y=count,
                              group=PLR, color=PLR)) +
  ylab("Phage count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(pro_lysischances$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_lysischances$PLR)+2))

phagecount_plot + facet_wrap(~PIV)
```

Phage int val
```{r}
phageintval_plot <- ggplot(data=pro_phagevals, aes(x=update, y=mean_intval,
                                               group=PLR, color=PLR)) +
  ylab("Phage Interaction value") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(pro_phagevals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_phagevals$PLR)+2))

phageintval_plot + facet_wrap(~PIV)
```

Chance of lysis
```{r}
pro_lysischances2 <- pro_lysischances %>%
  mutate(phage_type = as.factor(ifelse(PIV == "0", "A - Harmful Phage", 
                             ifelse(PIV == "0.5", "B - Neutral Phage", 
                                    "C - Beneficial Phage")))) %>%
  mutate(ProphageDeathRate = as.factor(ifelse(PLR == "0", "0",
                                              ifelse(PLR == "0.025", "0.025",
                                                     "0.05"))))

lysischances_plot <- ggplot(data=pro_lysischances2,
                            aes(x=update, y=mean_lysischance,
                                group=ProphageDeathRate, color=ProphageDeathRate)) +
  ylab("Mean Chance of Lysis") + xlab("Evolutionary Time (Updates)") + 
  stat_summary(aes(color=ProphageDeathRate, fill=ProphageDeathRate),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + 
  scale_colour_manual(values=plasma(nlevels(pro_lysischances2$ProphageDeathRate)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_lysischances2$ProphageDeathRate)+2))

lysischances_plot + facet_wrap(~phage_type)
```

Chance of Lysis Rainplot
```{r}
pro_final_lysischances2 <- pro_final_lysischances %>%
  mutate(phage_type = as.factor(ifelse(PIV == "0", "1 - Harmful Phage", 
                             ifelse(PIV == "0.5", "2 - Neutral Phage", 
                                    "3 - Beneficial Phage"))))

lysis_raincloud <- ggplot(data = pro_final_lysischances2, 
                              aes(x = PLR, y = mean_lysischance,
                                  group=PLR, color=PLR)) + 
  geom_boxplot(
    width = .25, 
    outlier.shape = NA
  ) +
  geom_point(
    size = 1.3,
    alpha = .3,
    position = position_jitter(
      seed = 1, width = .1
    )
  ) + 
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  ylab("Mean Chance of Lysis") + xlab("Prophage Loss Rate") +
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") +
  scale_colour_manual(values=plasma(5)) + 
  scale_fill_manual(values=plasma(5)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "none")

lysis_raincloud + facet_wrap(~phage_type)
```

Mean difference in incorporation vals
```{r}
incorp_diff_plot <- ggplot(data=pro_incvalsdiffs,
                            aes(x=update, y=mean_incval_difference),
                                group=PLR, color=PLR) +
  ylab("Mean difference in inc vals") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(pro_incvalsdiffs$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(pro_incvalsdiffs$PLR)+2))

incorp_diff_plot + facet_wrap(~PIV)
```

Difference in incorporation vals rainplot
```{r}
pro_final_incvalsdiffs2 <- pro_final_incvalsdiffs %>%
  mutate(phage_type = as.factor(ifelse(PIV == "0", "1 - Harmful Phage", 
                             ifelse(PIV == "0.5", "2 - Neutral Phage", 
                                    "3 - Beneficial Phage")))) %>%
  mutate(compatibility = 1 - mean_incval_difference)

incvals_raincloud <- ggplot(data = pro_final_incvalsdiffs2, 
                              aes(x = PLR, y = compatibility,
                                  group=PLR, color=PLR)) + 
  geom_boxplot(
    width = .25, 
    outlier.shape = NA
  ) +
  geom_point(
    size = 1.3,
    alpha = .3,
    position = position_jitter(
      seed = 1, width = .1
    )
  ) + 
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  ylab("Mean Host-Phage Compatibility") + xlab("Prophage Loss Rate") +
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") +
  scale_colour_manual(values=plasma(5)) + 
  scale_fill_manual(values=plasma(5)) +
  theme(legend.position = "none")

incvals_raincloud + facet_wrap(~phage_type)
```

Correlation between chance of lysis and difference in incorpation values
```{r}
pro_final_lysis_induction <- inner_join(pro_final_lysischances2, pro_final_incvalsdiffs2,
                                        by = c("SEED", "PLR", "PIV", "phage_type")) %>%
  select(mean_lysischance, mean_incval_difference, SEED, PLR, PIV, phage_type) %>%
  mutate(mean_compatibility = 1 - mean_incval_difference)

lysis_induction_plot <- ggplot(pro_final_lysis_induction, 
                               aes(x = mean_lysischance, y = mean_compatibility,
                                   group = PLR, color = PLR)) +
  geom_point() +
  facet_wrap(~phage_type) +
  scale_colour_manual(values=plasma(5)) + 
  scale_fill_manual(values=plasma(5)) +
  ylab("Mean Host-Phage Compatibility") + xlab("Mean Chance of Lysis") +
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle=90, vjust=0.5), 
        panel.spacing.x = unit(3, "mm"))

lysis_induction_plot
```

### Stacked Histograms

Chance of lysis stacked histogram
```{r}
lysischance_stackedhistogram <- ggplot(pro_lysis_histdata, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab("Count of Phage with Phenotype") + xlab("Evolutionary time (in updates)") +
  scale_fill_manual("Chance of Lysis\n Phenotypes",values=plasma(nlevels(pro_lysis_histdata$Histogram_bins)+2)) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill="none") + guides(fill = guide_legend())

lysischance_stackedhistogram
```

Difference in inc vals stacked histogram
```{r}
incorp_diff_stackedhistogram <- ggplot(pro_incorp_diff_histdata, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab("Count of Phage with Phenotype") + xlab("Evolutionary time (in updates)") +
  scale_fill_manual("Difference in \n Incorporation Vals",
                    values=plasma(nlevels(pro_incorp_diff_histdata$Histogram_bins)+2)) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill="none") + guides(fill = guide_legend())

incorp_diff_stackedhistogram
```

