---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Induction Only - Benefit Off
The following includes the R scripts that we used to generate our graphs for the data where induction was possible for lysogenic phage, but the potential benefit to the hosts of lysogenic phage was turned off.

## Gather Settings and Treatments

General settings
```{r}
ind_folder <- "Data/InductionOnly/" #Folder where datafiles are 
ind_treatments <- c("PLR", "COI") #Names of the treatments being tested - should match what is in filenames
```

Stacked histogram settings
```{r}
induction_separate_by <- "PLR" #Facet wrap induction stacked histogram by Prophage Loss Rate
#How induction stacked histogram bins should be collapsed and renamed
induction_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Very low induction)", 
          Hist_0.1 = "0 to 0.2 (Very low induction)",
          Hist_0.2 = "0.2 to 0.4 (Moderately low induction)",
          Hist_0.3 = "0.2 to 0.4 (Moderately low induction)",
          Hist_0.4 = "0.4 to 0.6 (Moderate induction)",
          Hist_0.5 = "0.4 to 0.6 (Moderate induction)",
          Hist_0.6 = "0.6 to 0.8 (Moderately high induction)",
          Hist_0.7 = "0.6 to 0.8 (Moderately high induction)",
          Hist_0.8 = "0.8 to 1.0 (Very high induction)",
          Hist_0.9 = "0.8 to 1.0 (Very high induction)")

lysis_separate_by <- "PLR" #Facet wrap lysis chance stacked histogram by Prophage Loss Rate
#How lysis chance stacked histogram bins should be collapsed and renamed
lysis_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Highly lysogenic)",
                          Hist_0.1 = "0 to 0.2 (Highly lysogenic)",
                          Hist_0.2 = "0.2 to 0.4 (Moderately lysogenic)",
                          Hist_0.3 = "0.2 to 0.4 (Moderately lysogenic)",
                          Hist_0.4 = "0.4 to 0.6 (Temperate)",
                          Hist_0.5 = "0.4 to 0.6 (Temperate)",
                          Hist_0.6 = "0.6 to 0.8 (Moderately lytic)",
                          Hist_0.7 = "0.6 to 0.8 (Moderately lytic)",
                          Hist_0.8 = "0.8 to 1.0 (Highly lytic)",
                          Hist_0.9 = "0.8 to 1.0 (Highly lytic)")
```

## Collect and Munge Data

Gather filenames
```{r}
ind_all_filenames <- list.files(ind_folder)

ind_freeliving_filenames <- str_subset(ind_all_filenames, "FreeLivingSyms")
ind_hostval_filenames <- str_subset(ind_all_filenames, "HostVals")
ind_induction_filenames <- str_subset(ind_all_filenames, "InductionChance")
ind_lysischance_filenames <- str_subset(ind_all_filenames, "LysisChance")
ind_phagevals_filenames <- str_subset(ind_all_filenames, "SymVals")
```

Combine time series data for all subsets of datafiles
```{r, results='hide', message=FALSE}
ind_freeliving <- combine_time_data(ind_freeliving_filenames, ind_folder, ind_treatments)
ind_hostvals <- combine_time_data(ind_hostval_filenames, ind_folder, ind_treatments)
ind_inductionchances <- combine_time_data(ind_induction_filenames, ind_folder, ind_treatments)
ind_lysischances <- combine_time_data(ind_lysischance_filenames, ind_folder, ind_treatments)
ind_phagevals <- combine_time_data(ind_phagevals_filenames, ind_folder, ind_treatments)
```

Rearrange time series data into stacked histogram data
```{r}
ind_induction_histdata <- combine_histogram_data(ind_inductionchances, 
                                             induction_separate_by,
                                             induction_histogram_bins)

ind_lysis_histdata <- combine_histogram_data(ind_lysischances, 
                                         lysis_separate_by, 
                                         lysis_histogram_bins)
```

Extract information about average genome values at the end of the runs
```{r}
ind_final_freeliving <- last_update(ind_freeliving)
ind_final_hostvals <- last_update(ind_hostvals)
ind_final_inductionchances <- last_update(ind_inductionchances)
ind_final_lysischances <- last_update(ind_lysischances)
ind_final_phagevals <- last_update(ind_phagevals)
```

## Analyze Data
### Summary Stats
```{r}
ind_lysis_summary_stats <- ind_final_lysischances %>%
  group_by(PLR, COI) %>%
  summarise(final_chance_of_lysis = mean(mean_lysischance*100))

ind_lysis_summary_stats
```

```{r}
ind_induction_summary_stats <- ind_final_inductionchances %>%
   group_by(PLR, COI) %>%
   summarise(Final_COI = mean(mean_inductionchance*100))

ind_induction_summary_stats
```

### Wilcox tests

0% chance of induction
```{r}
#At a chance of induction of 0%, compare PLR 0% to PLR 2.5%
ind_PLR0 <- subset(subset(ind_final_inductionchances, PLR == "0"), COI == "0")
ind_PLR0.025 <- subset(subset(ind_final_inductionchances, PLR == "0.025"), COI == "0")

wilcox.test(ind_PLR0$mean_inductionchance, ind_PLR0.025$mean_inductionchance)$p.value * bonferroni

#At a chance of induction of 0%, compare PLR 0% to PLR 5%
ind_PLR0.05 <- subset(subset(ind_final_inductionchances, PLR == "0.05"), COI == "0")

wilcox.test(ind_PLR0$mean_inductionchance, ind_PLR0.05$mean_inductionchance)$p.value * bonferroni
```

Compare Chance of Lysis for PLR 2.5% across all levels of induction
```{r}
noInd_PLR2.5 <- subset(plr_final_lysischances, PLR == "0.025")
ind_PLR2.5_COI0.0 <- subset(subset(ind_final_lysischances, PLR == "0.025"), COI == "0")
ind_PLR2.5_COI0.1 <- subset(subset(ind_final_lysischances, PLR == "0.025"), COI == "0.1")

wilcox.test(noInd_PLR2.5$mean_lysischance, ind_PLR2.5_COI0.0$mean_lysischance)$p.value * bonferroni
wilcox.test(noInd_PLR2.5$mean_lysischance, ind_PLR2.5_COI0.1$mean_lysischance)$p.value * bonferroni
```

Compare Chance of Lysis for PLR 5% across all levels of induction
```{r}
noInd_PLR5 <- subset(plr_final_lysischances, PLR == "0.05")
ind_PLR5_COI0.0 <- subset(subset(ind_final_lysischances, PLR == "0.05"), COI == "0")
ind_PLR5_COI0.1 <- subset(subset(ind_final_lysischances, PLR == "0.05"), COI == "0.1")

wilcox.test(noInd_PLR5$mean_lysischance, ind_PLR5_COI0.0$mean_lysischance)$p.value * bonferroni
wilcox.test(noInd_PLR5$mean_lysischance, ind_PLR5_COI0.1$mean_lysischance)$p.value * bonferroni
```

## Generate Graphs
### Host graphs
Host count
```{r}
hostcount_plot <- ggplot(data=ind_hostvals,
                         aes(x=update, y=count, 
                             group=PLR, colour=PLR)) + 
  ylab("Host Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(ind_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_hostvals$PLR)+2))

hostcount_plot + facet_wrap(~COI)
```

Host cfu count
```{r}
hostcount_cfu_plot <- ggplot(data=ind_hostvals, aes(x=update, y=cfu_count, 
                                                group=PLR, colour=PLR)) + 
  ylab("Host CFU Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(ind_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_hostvals$PLR)+2))

hostcount_cfu_plot + facet_wrap(~COI)
```

Host uninfected count
```{r}
host_uninfected_plot <- ggplot(data=ind_hostvals, aes(x=update, y=uninfected_host_count, 
                                                  group=PLR, colour=PLR)) + 
  ylab("Uninfected Host Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(ind_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_hostvals$PLR)+2))

host_uninfected_plot + facet_wrap(~COI)
```

Host int vals
```{r}
hostvals_plot <- ggplot(data=ind_hostvals, aes(x=update, y=mean_intval, 
                                           group=PLR, colour=PLR)) + 
  ylab("Mean Host Interaction Value") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(ind_hostvals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_hostvals$PLR)+2))

hostvals_plot + facet_wrap(~COI)
```

### Phage graphs

Phage count
```{r}
phagecount_plot <- ggplot(data=ind_lysischances,
                          aes(x=update, y=count,
                              group=PLR, color=PLR)) +
  ylab("Phage count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(ind_lysischances$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_lysischances$PLR)+2))

phagecount_plot + facet_wrap(~COI)
```

Phage int val
```{r}
phageintval_plot <- ggplot(data=ind_phagevals, aes(x=update, y=mean_intval,
                                               group=PLR, color=PLR)) +
  ylab("Phage Interaction value") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(ind_phagevals$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_phagevals$PLR)+2))

phageintval_plot + facet_wrap(~COI)
```

Chance of lysis
```{r}
lysischances_plot <- ggplot(data=ind_lysischances,
                            aes(x=update, y=mean_lysischance,
                                group=PLR, color=PLR)) +
  ylab("Mean Chance of Lysis") + xlab("Evolutionary Time (Updates)") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + 
  scale_colour_manual(values=plasma(5)) + 
  scale_fill_manual(values=plasma(5))

lysischances_plot + facet_wrap(~COI)
```
Chance of lysis rainplot
```{r}
ind_final_lysischances2 <- ind_final_lysischances %>%
  mutate(starting_ind = as.factor(ifelse(COI == "0",
                                         "Induction Chance 0%",
                                         "Induction Chance 10%")))

lysis_raincloud <- ggplot(data = ind_final_lysischances2, 
                              aes(x = PLR, y = mean_lysischance,
                                  group=PLR, color=PLR)) + 
  geom_boxplot(
    width = .25, 
    outlier.shape = NA
  ) +
  geom_point(
    size = 1.3,
    alpha = .3,
    position = position_jitter(
      seed = 1, width = .1
    )
  ) + 
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  ylab("Mean Chance of Lysis") + xlab("Prophage Loss Rate") +
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") +
  scale_colour_manual(values=plasma(5)) + 
  scale_fill_manual(values=plasma(5)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "none")

lysis_raincloud + facet_wrap(~starting_ind)
```

Chance of induction
```{r}
inductionchances_plot <- ggplot(data=ind_inductionchances,
                            aes(x=update, y=mean_inductionchance,
                                group=PLR, color=PLR)) +
  ylab("Mean chance of induction") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + scale_colour_manual(values=plasma(nlevels(ind_inductionchances$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_inductionchances$PLR)+2))

inductionchances_plot + facet_wrap(~COI)
```

Chance of induction - boxplot
```{r}
inductionchances_boxplot <- ggplot(data = ind_final_inductionchances,
                                   aes(y=mean_inductionchance,
                                       group=PLR, color=PLR)) +
  geom_boxplot() + facet_wrap(~COI) + ylab("Mean chance of induction") + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") + theme(axis.ticks.x = element_blank()) +
  scale_colour_manual(values=plasma(nlevels(ind_final_inductionchances$PLR)+2)) + 
  scale_fill_manual(values=plasma(nlevels(ind_final_inductionchances$PLR)+2))

inductionchances_boxplot
```

Chance of induction - raincloud plot
```{r}
ind_final_inductionchances2 <- ind_final_inductionchances %>%
  mutate(starting_ind = as.factor(ifelse(COI == "0",
                                         "Induction Chance 0%",
                                         "Induction Chance 10%")))

induction_raincloud <- ggplot(data = ind_final_inductionchances2, 
                              aes(x = PLR, y = mean_inductionchance,
                                  group=PLR, color=PLR)) + 
  geom_boxplot(
    width = .25, 
    outlier.shape = NA
  ) +
  geom_point(
    size = 1.3,
    alpha = .3,
    position = position_jitter(
      seed = 1, width = .1
    )
  ) + 
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  ylab("Mean Chance of Induction") + xlab("Prophage Loss Rate") +
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill="none") +
  scale_colour_manual(values=plasma(5)) + 
  scale_fill_manual(values=plasma(5)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "none")

induction_raincloud + facet_wrap(~starting_ind)
```

### Stacked Histograms

Chance of lysis stacked histogram
```{r}
lysischance_stackedhistogram <- ggplot(ind_lysis_histdata, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab("Count of Phage with Phenotype") + xlab("Evolutionary time (in updates)") +
  scale_fill_manual("Chance of Lysis\n Phenotypes",values=plasma(nlevels(ind_lysis_histdata$Histogram_bins)+2)) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill="none") + guides(fill = guide_legend())

lysischance_stackedhistogram
```

Chance of induction stacked histogram
```{r}
inductionchance_stackedhistogram <- ggplot(ind_induction_histdata, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab("Count of Phage with Phenotype") + xlab("Evolutionary time (in updates)") +
  scale_fill_manual("Chance of Induction\n Phenotypes",
                    values=plasma(nlevels(ind_induction_histdata$Histogram_bins)+2)) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill="none") + guides(fill = guide_legend())

inductionchance_stackedhistogram
```
