---
title: "Symbulation Graphs"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(readr)
library(dplyr)
library(stringr)
library(Hmisc)
library(tidyr)
library(forcats)
library(viridis)
```

# Load Functions

For a particular treatment, extract the value for the data from the filename using string manipulation
```{r}
get_treatment_val <- function(filename, treatment_name){
  #Input: a filename string and a string representing which treatment you are looking for
  #       Ex: "HostVals_VT0.6_SEED10.data" and "VT"
  #Output: A string representing the value for the given treatment
  #       Ex: "0.6"
  number_pattern <- "[\\d][._][\\d]*"
  treatment_pattern <- str_c(c(treatment_name, number_pattern), sep="", collapse="")
  treatment <- str_extract(filename, treatment_pattern)
  treatment_val <- NA
  
  decimal_pattern <- "[\\d].[\\d]+"
  integer_pattern <- "[\\d]+"
  if(str_detect(treatment, "\\.")){
    treatment_val <- str_extract(treatment, decimal_pattern)
  } else {
    treatment_val <- str_extract(treatment, integer_pattern)
  }
  treatment_val
}
```

Retrieve the dataset for a particular filename, adding columns for each treatment
```{r}
get_data <- function(filename, folder, treatments){
  #Input: A datafile name as a string, the folder where the data is located, and a list of treatment names
  #       Ex: "HostVals_VT0.6_SEED10.data", "VertTrans_Data/", and c("VT")
  #Output: A tibble containing all data in the file with a column added for the Seed and for all treatment names
  #       Ex: The columns added for the above data would be SEED with all values at 10 and 
  #           VT with all values at 0.6
  seed <- str_extract(filename, "SEED[\\d]+.")
  seed_val <- str_extract(seed, "[\\d]+")
  
  full_filename <- str_c(c(folder, filename), sep="", collapse="")
  data <- read_csv(full_filename) %>% mutate(SEED = seed_val)
  
  for (i in c(1:length(treatments))){
    data <- data %>% mutate(!!treatments[i] := as.factor(get_treatment_val(filename, treatments[i])))
  }
  data
}
```

For all files of a particular type (e.g. HostVals), create one dataset that can be used for time series graphs
```{r}
combine_time_data <- function(filenames, folder, treatments){
  #Input: A list of relevant datafile names, the folder where the data is, and a list of treatment names
  #       Ex: c("HostVals_VT0.6_SEED10.data", "HostVals_VT0.8_SEED10.data"), "VertTrans_Data/", 
  #           and c("VT")
  #Output: A tibble with all data from all datafiles given, with columns added for the seed and 
  #       all treatment names
  all_data <- get_data(filenames[1], folder, treatments)
  
  if(length(filenames) > 1){
    for (i in c(2:length(filenames))){
      add_data <- get_data(filenames[i], folder, treatments)
      all_data <- full_join(all_data, add_data)
    }
  }
  all_data %>% mutate(SEED = as.factor(SEED))
}
```

Take the time series data and turn it into a format that can be used for stacked histograms
```{r}
combine_histogram_data <- function(time_data, separate_by, bins){
  #Input: The time series data produced by combine_time_data, the column name 
  #       which the stacked histogram will be facet wrapped by,
  #       and 1:1 list for the desired names of histogram bins
  #       Ex: lysischances, "PLR", and
  #           lysis_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Highly lysogenic)",
                          # Hist_0.1 = "0 to 0.2 (Highly lysogenic)",
                          # Hist_0.2 = "0.2 to 0.4 (Moderately lysogenic)",
                          # Hist_0.3 = "0.2 to 0.4 (Moderately lysogenic)",
                          # Hist_0.4 = "0.4 to 0.6 (Temperate)",
                          # Hist_0.5 = "0.4 to 0.6 (Temperate)",
                          # Hist_0.6 = "0.6 to 0.8 (Moderately lytic)",
                          # Hist_0.7 = "0.6 to 0.8 (Moderately lytic)",
                          # Hist_0.8 = "0.8 to 1.0 (Highly lytic)",
                          # Hist_0.9 = "0.8 to 1.0 (Highly lytic)")
  #Output: A tibble with 4 columns: update, treatment, Histogram_bins, and count. 
  #       This is the format needed to create a stacked histogram.
  
  #collect initial data and pivot longer
  initial_data <- time_data %>%
    pivot_longer(
      cols = starts_with("Hist_"),
      names_to = "Histogram_bin",
      values_to = "Bin_count"
    ) %>%
    mutate(Histogram_bin = as.factor(Histogram_bin))
  
  #collapse the bins
  collapsed_data <- initial_data
  collapsed_data$Histogram_factor <- dplyr::recode(collapsed_data$Histogram_bin, !!!bins)
  
  #aggregate the counts
  collapsed_data <- collapsed_data %>%
    mutate(count = Bin_count,
           treatment = get(separate_by)) %>%
    select(treatment, update, Histogram_factor, count)
  
  final_data <- aggregate(list(count = collapsed_data$count), 
                          list(update = collapsed_data$update,
                               treatment = collapsed_data$treatment,
                               Histogram_bins = collapsed_data$Histogram_factor), sum)
  final_data
}
```

# Gather Settings and Treatments

General settings
```{r}
folder <- "Data/" #Folder where datafiles are 
treatments <- c("PLR", "PIV") #Names of the treatments being tested - should match what is in filenames
```

Stacked histogram settings
```{r}
induction_separate_by <- "PLR" #Facet wrap induction stacked histogram by Prophage Loss Rate
#How induction stacked histogram bins should be collapsed and renamed
induction_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Very low induction)", 
          Hist_0.1 = "0 to 0.2 (Very low induction)",
          Hist_0.2 = "0.2 to 0.4 (Moderately low induction)",
          Hist_0.3 = "0.2 to 0.4 (Moderately low induction)",
          Hist_0.4 = "0.4 to 0.6 (Moderate induction)",
          Hist_0.5 = "0.4 to 0.6 (Moderate induction)",
          Hist_0.6 = "0.6 to 0.8 (Moderately high induction)",
          Hist_0.7 = "0.6 to 0.8 (Moderately high induction)",
          Hist_0.8 = "0.8 to 1.0 (Very high induction)",
          Hist_0.9 = "0.8 to 1.0 (Very high induction)")

lysis_separate_by <- "PLR" #Facet wrap lysis chance stacked histogram by Prophage Loss Rate
#How lysis chance stacked histogram bins should be collapsed and renamed
lysis_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Highly lysogenic)",
                          Hist_0.1 = "0 to 0.2 (Highly lysogenic)",
                          Hist_0.2 = "0.2 to 0.4 (Moderately lysogenic)",
                          Hist_0.3 = "0.2 to 0.4 (Moderately lysogenic)",
                          Hist_0.4 = "0.4 to 0.6 (Temperate)",
                          Hist_0.5 = "0.4 to 0.6 (Temperate)",
                          Hist_0.6 = "0.6 to 0.8 (Moderately lytic)",
                          Hist_0.7 = "0.6 to 0.8 (Moderately lytic)",
                          Hist_0.8 = "0.8 to 1.0 (Highly lytic)",
                          Hist_0.9 = "0.8 to 1.0 (Highly lytic)")

incorp_diff_separate_by <- "PIV"
incorp_diff_histogram_bins <- c(Hist_0.0 = "0 to 0.2 (Highly beneficial)",
                          Hist_0.1 = "0 to 0.2 (Highly beneficial)",
                          Hist_0.2 = "0.2 to 0.4 (Moderately beneficial)",
                          Hist_0.3 = "0.2 to 0.4 (Moderately beneficial)",
                          Hist_0.4 = "0.4 to 0.6 (Neutral)",
                          Hist_0.5 = "0.4 to 0.6 (Neutral)",
                          Hist_0.6 = "0.6 to 0.8 (Moderately detrimental)",
                          Hist_0.7 = "0.6 to 0.8 (Moderately detrimental)",
                          Hist_0.8 = "0.8 to 1.0 (Highly detrimental)",
                          Hist_0.9 = "0.8 to 1.0 (Highly detrimental)")
```

# Collect and Munge Data

Gather filenames
```{r}
all_filenames <- list.files(folder)

bacteriumincvals_filenames <- str_subset(all_filenames, "BacteriumIncVals")
freeliving_filenames <- str_subset(all_filenames, "FreeLivingSyms")
hostval_filenames <- str_subset(all_filenames, "HostVals")
incvaldiffs_filenames <- str_subset(all_filenames, "IncValDifferences")
induction_filenames <- str_subset(all_filenames, "InductionChance")
lysischance_filenames <- str_subset(all_filenames, "LysisChance")
phageincvals_filenames <- str_subset(all_filenames, "PhageIncVals")
phagevals_filenames <- str_subset(all_filenames, "SymVals")
```

Combine time series data for all subsets of datafiles
```{r, results='hide', message=FALSE}
bacteriumincvals <- combine_time_data(bacteriumincvals_filenames, folder, treatments)
freeliving <- combine_time_data(freeliving_filenames, folder, treatments)
hostvals <- combine_time_data(hostval_filenames, folder, treatments)
incvalsdiffs <- combine_time_data(incvaldiffs_filenames, folder, treatments)
inductionchances <- combine_time_data(induction_filenames, folder, treatments)
lysischances <- combine_time_data(lysischance_filenames, folder, treatments)
phageincvals <- combine_time_data(phageincvals_filenames, folder, treatments)
phagevals <- combine_time_data(phagevals_filenames, folder, treatments)
```

Rearrange time series data into stacked histogram data
```{r}
induction_histdata <- combine_histogram_data(inductionchances, 
                                             induction_separate_by,
                                             induction_histogram_bins)

lysis_histdata <- combine_histogram_data(lysischances, 
                                         lysis_separate_by, 
                                         lysis_histogram_bins)

incorp_diff_histdata <- combine_histogram_data(incvalsdiffs,
                                                     incorp_diff_separate_by,
                                                     incorp_diff_histogram_bins)
```

# Create Graphs

## Templates
Time Series template
```{r, eval=FALSE}
#All values with this format: <VAL_NAME> need to be filled in
timeseries_graph <- ggplot(data= <TIME_SERIES_DATASET_NAME>,
                         aes(x=update, y= <Y_AXIS_VAL>, 
                             group=<GROUP_VAL_NAME>, colour=<GROUP_VAL_NAME>)) + 
  ylab(<Y_AXIS_LABEL>) + xlab("Updates") + 
  stat_summary(aes(color=<GROUP_VAL_NAME>, fill=<GROUP_VAL_NAME>),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + 
  scale_colour_manual(values=viridis(nlevels(<TIME_SERIES_DATASET_NAME>$<GROUP_VAL_NAME>))) + 
  scale_fill_manual(values=viridis(nlevels(<TIME_SERIES_DATASET_NAME>$<GROUP_VAL_NAME>)))

#Facet wrapping is optional, but can be used if multiple treatments are being used
timeseries_graph + facet_wrap(<GROUP_VAL2_NAME>) 
```

Stacked Histogram template
```{r, eval=FALSE}
#All values with this format: <VAL_NAME> need to be filled in
stackedhistogram <- ggplot(data = <HISTOGRAM_DATASET_NAME>, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab(<Y_AXIS_LABEL>) + xlab("Evolutionary time (in updates)") +
  scale_fill_manual(<LEGEND_TITLE>,values=viridis(nlevels(<HISTOGRAM_DATASET_NAME>$Histogram_bins))) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill=FALSE) + guides(fill = guide_legend())

stackedhistogram
```

## Lysis vs. Lysogeny

### Host graphs
Host count
```{r}
hostcount_plot <- ggplot(data=hostvals,
                         aes(x=update, y=count, 
                             group=PLR, colour=PLR)) + 
  ylab("Host Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(hostvals$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(hostvals$PLR)))

hostcount_plot + facet_wrap(~PIV)
```

Host cfu count
```{r}
hostcount_cfu_plot <- ggplot(data=hostvals, aes(x=update, y=cfu_count, 
                                                group=PLR, colour=PLR)) + 
  ylab("Host CFU Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(hostvals$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(hostvals$PLR)))

hostcount_cfu_plot + facet_wrap(~PIV)
```

Host uninfected count
```{r}
host_uninfected_plot <- ggplot(data=hostvals, aes(x=update, y=uninfected_host_count, 
                                                  group=PLR, colour=PLR)) + 
  ylab("Uninfected Host Count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(hostvals$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(hostvals$PLR)))

host_uninfected_plot + facet_wrap(~PIV)
```

Host int vals
```{r}
hostvals_plot <- ggplot(data=hostvals, aes(x=update, y=mean_intval, 
                                           group=PLR, colour=PLR)) + 
  ylab("Mean Host Interaction Value") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(hostvals$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(hostvals$PLR)))

hostvals_plot + facet_wrap(~PIV)
```

### Phage graphs

Phage count
```{r}
phagecount_plot <- ggplot(data=lysischances,
                          aes(x=update, y=count,
                              group=PLR, color=PLR)) +
  ylab("Phage count") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(lysischances$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(lysischances$PLR)))

phagecount_plot + facet_wrap(~PIV)
```

Phage int val
```{r}
phageintval_plot <- ggplot(data=phagevals, aes(x=update, y=mean_intval,
                                               group=PLR, color=PLR)) +
  ylab("Phage Interaction value") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(phagevals$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(phagevals$PLR)))

phageintval_plot + facet_wrap(~PIV)
```

Chance of lysis
```{r}
lysischances_plot <- ggplot(data=lysischances,
                            aes(x=update, y=mean_lysischance,
                                group=PLR, color=PLR)) +
  ylab("Mean chance of lysis") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(lysischances$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(lysischances$PLR)))

lysischances_plot + facet_wrap(~PIV)
```

Chance of induction
```{r}
inductionchances_plot <- ggplot(data=inductionchances,
                            aes(x=update, y=mean_inductionchance,
                                group=PLR, color=PLR)) +
  ylab("Mean chance of induction") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(inductionchances$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(inductionchances$PLR)))

inductionchances_plot + facet_wrap(~PIV)
```

Mean difference in incorporation vals
```{r}
incorp_diff_plot <- ggplot(data=incvalsdiffs,
                            aes(x=update, y=mean_incval_difference),
                                group=PLR, color=PLR) +
  ylab("Mean difference in inc vals") + xlab("Updates") + 
  stat_summary(aes(color=PLR, fill=PLR),
               fun.data="mean_cl_boot", geom=c("smooth"), se=TRUE) + 
  theme(panel.background = element_rect(fill='white', colour='black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  guides(fill=FALSE) + scale_colour_manual(values=viridis(nlevels(inductionchances$PLR))) + 
  scale_fill_manual(values=viridis(nlevels(inductionchances$PLR)))

incorp_diff_plot + facet_wrap(~PIV)
```

Chance of lysis stacked histogram
```{r}
lysischance_stackedhistogram <- ggplot(lysis_histdata, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab("Count of Phage with Phenotype") + xlab("Evolutionary time (in updates)") +
  scale_fill_manual("Chance of Lysis\n Phenotypes",values=viridis(nlevels(lysis_histdata$Histogram_bins))) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill=FALSE) + guides(fill = guide_legend())

lysischance_stackedhistogram
```

Chance of induction stacked histogram
```{r}
inductionchance_stackedhistogram <- ggplot(induction_histdata, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab("Count of Phage with Phenotype") + xlab("Evolutionary time (in updates)") +
  scale_fill_manual("Chance of Induction\n Phenotypes",
                    values=viridis(nlevels(induction_histdata$Histogram_bins))) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill=FALSE) + guides(fill = guide_legend())

inductionchance_stackedhistogram
```

Difference in inc vals stacked histogram
```{r}
incorp_diff_stackedhistogram <- ggplot(incorp_diff_histdata, 
                                       aes(update, count)) + 
  geom_area(aes(fill=Histogram_bins), position='stack') +
  ylab("Count of Phage with Phenotype") + xlab("Evolutionary time (in updates)") +
  scale_fill_manual("Difference in \n Incorporation Vals",
                    values=viridis(nlevels(incorp_diff_histdata$Histogram_bins))) +
  facet_wrap(~treatment) + 
  theme(panel.background = element_rect(fill='light grey', colour='black')) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  guides(fill=FALSE) + guides(fill = guide_legend())

incorp_diff_stackedhistogram
```
